---
title: "NMDSplot"
output: word_document
editor_options:
  chunk_output_type: console
---
```{r}
library(devtools)
#install_github("vqv/ggbiplot")
library(ggbiplot)
#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("sinhrks/ggfortify")
library(ggfortify)
#install.packages("FactoMineR")
library(FactoMineR)
library("ggfortify")
library(ggfortify)
library(rafalib)
#install.packages("ggplot2")
#install.packages("tidyverse")
library(ggplot2)
#install.packages("plotly")
library(plotly)
library(FactoMineR)
#install.packages("factoextra")
library(factoextra)
library(dplyr)
#install.packages('ggrepel')
library(ggrepel)
#install.packages("vegan")
library(vegan)
```

```{r}
setwd("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS")
getwd()

rc <- read.csv2("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/Salmonella_invitro-invivo-switch_raw_rc.csv", row.names=1)
View(rc)
class(rc)
str(rc)
head(rc)

me <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/meta_all.csv", sep=";")
View(me)
```

```{r}
library(DESeq2)
me$Group <- factor(me$Group)
d <- DESeqDataSetFromMatrix(countData=rc, colData=me, design=~1)
#View(d)
```
# Size factors

The first step is estimating size factors. The data is normalised for sequencing depth and compositional bias as done for the VST step. DESeq2 uses a method called *median-of-ratios* for this step.

```{r,block.title=TRUE}
d <- DESeq2::estimateSizeFactors(d,type="ratio")
#View(d)
#?estimateSizeFactors
class(d)
```
#Write out the Size factor normalized read counts!!!! :-)
```{r,block.title=TRUE}
dds <- estimateSizeFactors(d)
norm_data <- counts(dds, normalized=TRUE)
View(norm_data)

write.csv(norm_data,"D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/SF_all_1.csv",row.names=T)
```
#NMDS
```{r}
SF_norm <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/SF_all_1.csv",row.names=1)
View(SF_norm)
SF_norm_f <- SF_norm[rowSums(SF_norm)>0, ] #delete rows that all 0
View(SF_norm_f)
SF_norm_t <- t(SF_norm_f) #transform
View(SF_norm_t)
#make community matrix - extract columns with abundance information
matrix = SF_norm_t[,2:ncol(SF_norm_t)]
#turn abundance data frame into a matrix
df_matrix = as.matrix(matrix)
set.seed(10)
nmds = metaMDS(df_matrix, distance = "bray")
plot(nmds)
```

```{r}
#extract NMDS scores (x and y coordinates)
data.scores = as.data.frame(scores(nmds))
#add columns to data frame 
data.scores$SampleName = me$SampleName
data.scores$Model = me$Model
data.scores$Group = me$Group
 
head(data.scores)
```
data.scores$Sample = pc$Sample
data.scores$Time = pc$Time
data.scores$Type = pc$Type
```{r}
library(ggplot2)

NMDS_nice = ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) + 
    geom_point(size = 4, aes( shape = Group, colour = Model))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 12), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 14), 
    axis.title.x = element_text(face = "bold", size = 14, colour = "black"), 
    legend.title = element_text(size = 14, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "NMDS1", colour = "Model", y = "NMDS2", shape = "Group")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00","#005f9e")) 
NMDS_nice

```
###Labelling
```{r}
library(ggplot2)

NMDS_nice = ggplot(data.scores, aes(x = NMDS1, y = NMDS2),label = rownames(data.scores)) + 
    geom_point(size = 4, aes( shape = Group, colour = Model, ))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 12), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 14), 
    axis.title.x = element_text(face = "bold", size = 14, colour = "black"), 
    legend.title = element_text(size = 14, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "NMDS1", colour = "Model", y = "NMDS2", shape = "Group")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00","#005f9e")) +
    geom_text(aes(label = SampleName,fontface=2, hjust=0, vjust=-0.5))
NMDS_nice

```
##Filtering data for NMDS --> In vivo+early
```{r}
SF_norm <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/SF_all_1.csv",row.names=1)
View(SF_norm)
me <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/meta_all.csv", sep=";")
View(me)
me_e <- (me[c(1:39,49:51,63:68),])
View(me_e)
SF_norm_t <- t(SF_norm) #transform
View(SF_norm_t)
SF_norm_e <- (SF_norm_t[c(1:39,49:51,63:68),]) #filtering for early and switch
View(SF_norm_e)
SF_norm_f <- SF_norm_e[,colSums(SF_norm_e)>0] #delete columns that all 0
View(SF_norm_f)

#make community matrix - extract columns with abundance information
matrix = SF_norm_f[,2:ncol(SF_norm_f)]
#turn abundance data frame into a matrix
df_matrix = as.matrix(matrix)
set.seed(1000)
nmds = metaMDS(df_matrix, distance = "bray")
plot(nmds)

#extract NMDS scores (x and y coordinates)
data.scores = as.data.frame(scores(nmds))
#add columns to data frame 
data.scores$SampleName = me_e$SampleName
data.scores$Model = me_e$Model
data.scores$Group = me_e$Group
 
head(data.scores)
```

```{r}
library(ggplot2)

NMDS_nice = ggplot(data.scores, aes(x = NMDS1, y = NMDS2)) + 
    geom_point(size = 4, aes( shape = Group, colour = Model))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 12), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 14), 
    axis.title.x = element_text(face = "bold", size = 14, colour = "black"), 
    legend.title = element_text(size = 14, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "NMDS1", colour = "Model", y = "NMDS2", shape = "Group")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00","#005f9e")) 
NMDS_nice

```

```{r}
library(ggplot2)

NMDS_nice = ggplot(data.scores, aes(x = NMDS1, y = NMDS2),label = rownames(data.scores)) + 
    geom_point(size = 4, aes( shape = Group, colour = Model, ))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 12), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 14), 
    axis.title.x = element_text(face = "bold", size = 14, colour = "black"), 
    legend.title = element_text(size = 14, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "NMDS1", colour = "Model", y = "NMDS2", shape = "Group")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00","#005f9e")) +
    geom_text(aes(label = SampleName,fontface=2, hjust=0, vjust=-0.5))
NMDS_nice

```
##Filtering data for NMDS --> In vivo+early-CE
```{r}
SF_norm <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/SF_all_1.csv",row.names=1)
View(SF_norm)
me <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/meta_all.csv", sep=";")
View(me)
me_e <- (me[c(1:39,63:68),])
View(me_e)
SF_norm_t <- t(SF_norm) #transform
View(SF_norm_t)
SF_norm_e <- (SF_norm_t[c(1:39,63:68),]) #filtering for early and switch
View(SF_norm_e)
SF_norm_f <- SF_norm_e[,colSums(SF_norm_e)>0] #delete columns that all 0
View(SF_norm_f)

#make community matrix - extract columns with abundance information
matrix = SF_norm_f[,2:ncol(SF_norm_f)]
#turn abundance data frame into a matrix
df_matrix = as.matrix(matrix)
set.seed(1000)
nmds = metaMDS(df_matrix, distance = "bray")
plot(nmds)

#extract NMDS scores (x and y coordinates)
data.scores = as.data.frame(scores(nmds))
#add columns to data frame 
data.scores$SampleName = me_e$SampleName
data.scores$Model = me_e$Model
data.scores$Group = me_e$Group
 
head(data.scores)
```

```{r}
library(ggplot2)

NMDS_nice = ggplot(data.scores, aes(x = NMDS1, y = NMDS2),label = rownames(data.scores)) + 
    geom_point(size = 4, aes( shape = Group, colour = Model, ))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 12), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 14), 
    axis.title.x = element_text(face = "bold", size = 14, colour = "black"), 
    legend.title = element_text(size = 14, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "NMDS1", colour = "Model", y = "NMDS2", shape = "Group")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00","#005f9e")) +
    geom_text(aes(label = SampleName,fontface=2, hjust=0, vjust=-0.5))
NMDS_nice

```
##Filtering data for NMDS --> In vivo early CE-CC
```{r}
SF_norm <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/SF_all_1.csv",row.names=1)
#View(SF_norm)
me <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/meta_all.csv", sep=";")
View(me)
me_e <- (me[c(37:39,49),])
View(me_e)
SF_norm_t <- t(SF_norm) #transform
#View(SF_norm_t)
SF_norm_e <- (SF_norm_t[c(37:39,49),]) #filtering for early and switch
#View(SF_norm_e)
SF_norm_f <- SF_norm_e[,colSums(SF_norm_e)>0] #delete columns that all 0
#View(SF_norm_f)

#make community matrix - extract columns with abundance information
matrix = SF_norm_f[,2:ncol(SF_norm_f)]
#turn abundance data frame into a matrix
df_matrix = as.matrix(matrix)
set.seed(1000)
nmds = metaMDS(df_matrix, distance = "bray")
plot(nmds)

#extract NMDS scores (x and y coordinates)
data.scores = as.data.frame(scores(nmds))
#add columns to data frame 
data.scores$SampleName = me_e$SampleName
data.scores$Model = me_e$Model
data.scores$Group = me_e$Group
 
head(data.scores)
```

```{r}
library(ggplot2)

NMDS_nice = ggplot(data.scores, aes(x = NMDS1, y = NMDS2),label = rownames(data.scores)) + 
    geom_point(size = 4, aes( shape = Group, colour = Model, ))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 12), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 14), 
    axis.title.x = element_text(face = "bold", size = 14, colour = "black"), 
    legend.title = element_text(size = 14, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "NMDS1", colour = "Model", y = "NMDS2", shape = "Group")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00","#005f9e")) +
    geom_text(aes(label = SampleName,fontface=2, hjust=0, vjust=-0.5))
NMDS_nice

```
##Filtering data for NMDS --> In vivo late
```{r}
SF_norm <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/SF_all_1.csv",row.names=1)
#View(SF_norm)
me <- read.csv("D:/RNAseq/preprocessing/data/in-vivo/PCA_NMDS/meta_all.csv", sep=";")
View(me)
me_e <- (me[c(37:39,63:68),])
View(me_e)
SF_norm_t <- t(SF_norm) #transform
#View(SF_norm_t)
SF_norm_e <- (SF_norm_t[c(37:39,63:68),]) #filtering for early and switch
#View(SF_norm_e)
SF_norm_f <- SF_norm_e[,colSums(SF_norm_e)>0] #delete columns that all 0
#View(SF_norm_f)

#make community matrix - extract columns with abundance information
matrix = SF_norm_f[,2:ncol(SF_norm_f)]
#turn abundance data frame into a matrix
df_matrix = as.matrix(matrix)
set.seed(1000)
nmds = metaMDS(df_matrix, distance = "bray", k=2, trymax = 20)
plot(nmds)


#extract NMDS scores (x and y coordinates)
data.scores = as.data.frame(scores(nmds))
#add columns to data frame 
data.scores$SampleName = me_e$SampleName
data.scores$Model = me_e$Model
data.scores$Group = me_e$Group
 
head(data.scores)
```

```{r}
library(ggplot2)

NMDS_nice = ggplot(data.scores, aes(x = NMDS1, y = NMDS2),label = rownames(data.scores)) + 
    geom_point(size = 4, aes( shape = Group, colour = Model, ))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 12), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 14), 
    axis.title.x = element_text(face = "bold", size = 14, colour = "black"), 
    legend.title = element_text(size = 14, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "NMDS1", colour = "Model", y = "NMDS2", shape = "Group")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00","#005f9e")) +
    geom_text(aes(label = SampleName,fontface=2, hjust=0, vjust=-0.5))
NMDS_nice

```